---
title: "SCP_TP53"
author: "Xiaoqian Liu"
date: "2025-2-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library("readxl")
library(glmnet)
library(ggplot2)
library(survival)
library(survminer)
library(factoextra)
# library(patchwork)
# library(ggnewscale)

source('~/Code/cluster_label.R')
source('~/Code/fused_Dist.R')
source('~/Code/fused_Lasso.R')
source('~/Code/fused_MCP.R')
source('~/Code/fused_SCAD.R')
source('~/Code/myGetDkn.R')
source('~/Code/pcox_path.R')
source('~/Code/pcox.R')
source('~/Code/pll_cox.R')
source('~/Code/wz.R')
source('~/Code/wRankAveg.R')
source('~/Code/trans_SCP.R')
source('~/Code/SCP.R')

```

## The TP53 database

The TP53 database has a collection of germline TP53 mutations with family history and age of diagnosis available, see <https://tp53.cancer.gov/get_tp53data#get_germ_mut> (germline TP53 variants and family history). Our current plan is to cluster germline TP53 mutations based on their survival effects on LFS patients. Since LFS is rare, we use all germline carriers information as the source data. 

```{r}

dt <- read.csv('GermlineDownload_r20.csv')

# select the needed columns
dt_use <- dt %>% select(Individual_code, Class, Germline_mutation, ProtDescription, Sex, Germline_carrier, Age, Age_at_diagnosis, Short_topo)
```


## LFS patients

To run the SCP model, we need the following information:

1) Mutation information 
2) Gender information
3) Survival information (either age or age at diagnosis is availabel)

Next, we extract the data from the TP53 database. Here note that we do not need to screen out those non-recurring patients since we will later borrow information from the general germline carriers to improve the estimation of survival effects.

```{r}
# select LFS patients with confirmed TP53 mutations
dt_LFS <- dt_use %>% filter(Class%in%c('LFS')&Germline_carrier%in% c('Confirmed', 'Obligatory'))
# dt_LFS <- dt_use %>% filter(Class%in%c('LFS'))
# screen out patients with no age information
dt_LFS <- dt_LFS %>% filter(!is.na(Age) | !is.na(Age_at_diagnosis))
# remove unknown mutations
dt_LFS <- dt_LFS %>% filter(ProtDescription!='p.?')

# create two new columns to include the survival information
dt_LFS <- dt_LFS %>% mutate(time=ifelse(is.na(Age_at_diagnosis), Age, Age_at_diagnosis))%>%
                            mutate(status=ifelse(is.na(Age_at_diagnosis), 0, 1))

# there are patients who have no gender information
dt_LFS <- dt_LFS %>% filter(!is.na(Sex))
# there is one patient who has a zero survival time
dt_LFS <- dt_LFS %>% filter(time>0)

# !!!!!! get the first cancer diagnosis information 
dt_LFS_first <- dt_LFS %>%
  group_by(Individual_code) %>%
  slice_min(Age_at_diagnosis, with_ties = FALSE) %>%
  ungroup()

# check the patient count for each mutation
mut_ct <- dt_LFS_first %>% group_by(ProtDescription)%>%summarise(n=n())
# only use recurring mutations
dt_LFS_first <- dt_LFS_first %>% filter(ProtDescription %in% mut_ct$ProtDescription[mut_ct$n>1])


# number of patients
nrow(dt_LFS_first)

# number of mutations
length(unique(dt_LFS_first$ProtDescription))
```

## Germline carriers

We extract the germline carriers (not LFS) from the TP53 database with the same set of mutations.

```{r}
# select LFS patients with confirmed TP53 mutations
dt_Germ <- dt_use %>% filter(!Class%in%c('LFS') & Germline_carrier!='negative'  | Class%in%c('LFS') &       
                               is.na(Germline_carrier) | Class%in%c('LFS')&Germline_carrier%in% c('50%prob.', ''))
# dt_Germ <- dt_use %>% filter(!Class%in%c('LFS') & Germline_carrier!='negative')

# screen out patients with no age information
dt_Germ <- dt_Germ %>% filter(!is.na(Age) | !is.na(Age_at_diagnosis))

# remove unknown mutations
dt_Germ <- dt_Germ %>% filter(ProtDescription!='p.?')

# create two new columns to include the survival information
dt_Germ <- dt_Germ %>% mutate(time=ifelse(is.na(Age_at_diagnosis), Age, Age_at_diagnosis))%>%
                            mutate(status=ifelse(is.na(Age_at_diagnosis), 0, 1))

# there are patients who have no gender information
dt_Germ <- dt_Germ %>% filter(!is.na(Sex))
# there is one patient who has a zero survival time
dt_Germ <- dt_Germ %>% filter(time>0)


# check the patient count for each mutation
# mut_ct <- dt_LFS_use %>% group_by(ProtDescription)%>%summarise(n=n())
# only use recurring mutations
# dt_LFS_use <- dt_LFS_use %>% filter(ProtDescription %in% mut_ct$ProtDescription[mut_ct$n>3])

# !!!!!! get the first cancer diagnosis information 
dt_Germ_first <- dt_Germ %>%
  group_by(Individual_code) %>%
  slice_min(Age_at_diagnosis, with_ties = FALSE) %>%
  ungroup()

# number of patients
nrow(dt_Germ_first)

# number of mutations
length(unique(dt_Germ_first$ProtDescription))
```

We only focus on those mutations which have carriers in both datasets. In that case, we can borrow the survival information from the general germline group to the LFS group. 

```{r}
mut_set <- intersect(unique(dt_LFS_first$ProtDescription), unique(dt_Germ_first$ProtDescription))

dt_LFS_use <- dt_LFS_first %>% filter(ProtDescription %in% mut_set)
dt_Germ_use <- dt_Germ_first %>% filter(ProtDescription %in% mut_set)

length(mut_set)
dim(dt_LFS_use)
dim(dt_Germ_use)

table(dt_LFS_use$ProtDescription)
table(dt_Germ_use$ProtDescription)
```

## Cox-kmeans for LFS only

We first run Cox-kmeans by only using the LFS patients data.
```{r}
n_T <- nrow(dt_LFS_use)
# mutation list
mut_list <- unique(dt_LFS_use$ProtDescription)
### Covariate matrix
X_T <- matrix(0, nrow = n_T, ncol = length(mut_list)+1)
colnames(X_T) <- c('gender', mut_list)
# covariates include gender + mutations
for (i in 1:n_T) {
  # put gender into the first column
  X_T[i, 1] <-  ifelse(dt_LFS_use$Sex[i]=='M', 1, 0) # gender. Male is 1, female is 0 
  # get the corresponding mutations 
  ind <- which(colnames(X_T)==dt_LFS_use$ProtDescription[i])
  X_T[i, ind] <- 1
}

### Survival information
Y_T <- cbind(time = dt_LFS_use$time, status = dt_LFS_use$status) 

## Preliminary estimation and ranking for constructing D
fit <- glmnet(X_T, Y_T, family = "cox", alpha = 0, lambda = 0)

k_select <- fviz_nbclust(as.matrix(fit$beta), kmeans, method = "silhouette")
cluster.cox <- cluster_label_kmeans(fit$beta, which.max(k_select$data$y))
llk_cox <- pll_cox(Y_T[, 1], Y_T[, 2], X_T, fit$beta)

#nu_cox <-  length(which(abs(D%*%out$beta)>1e-2))
bic_cox <- -2*llk_cox + log(nrow(X_T))*(ncol(X_T)-1)
  
```


## SCP for LFS only

We first run SCP by only using the LFS patients data.

```{r}
## Construct the D matrix
  beta0 <- as.vector(fit$beta[-1]) # we do not include the gender
  # sort X and then use the first-order differencing matrix
  ss <- sort(beta0, index.return = TRUE)
  tau <- ss$ix
  X1 <- X_T[, -1]
  X1_order <- X1[, tau]
  X_order <- cbind(X_T[, 1], X1_order)
  D <- myGetDkn(k=1, n=ncol(X1_order))
  D <- cbind(rep(0, nrow(D)), D)


  # Loop
ncSeq <- 1:5
nC <- length(ncSeq)
llkSeq <- rep(NA, nC)
bicSeq <- rep(NA, nC)
aicSeq <- rep(NA, nC)
nuSeq <- rep(NA, nC)
Beta_matrix <- matrix(NA, ncol(X_T), nC)

for (j in 1:nC) {
  cat(paste('Current stage:', j, 'case'))
  # fix the number of clusters
  nclusters <- ncSeq[j]
  
  
  t0 <- Sys.time()
  out <- pcox(Y_T[, 1], Y_T[, 2], X_order, D, penalty="distance",lambda=0, k=nclusters-1, 
                                    #beta0 = rnorm(ncol(X_order), 0,1),
                                     #beta0 = c(fit$beta[1], sort(beta0)),
                                    maxiters=1e3, tol=1e-5)
  t1 <- Sys.time()
  t1-t0

  beta_hat <- beta_order <- out$beta[-1]
  for(i in 1:(length(beta_order))){
    beta_hat[tau[i]] <- beta_order[i]
  }
  beta_hat <- c(out$beta[1], beta_hat)
  Beta_matrix[, j] <- beta_hat 

  
  llkSeq[j] <- pll_cox(Y_T[, 1], Y_T[, 2], X_T, beta_hat)

  nu0 <- nuSeq[j] <-  length(which(abs(D%*%out$beta)>1e-2))
  #tau0 <- choose(nrow(D), nu0)
  bicSeq[j] <- -2*llkSeq[j] + log(nrow(X_order))*nu0
  aicSeq[j] <- -2*llkSeq[j] + 2*nu0
  
}

save.image("TP53_test_final.RData")
```

```{r}
idx <- which.min(bicSeq)
st <- sort(Beta_matrix[-1, idx], index.return = TRUE)

plot(st$x)

which(diff(st$x)>0.05)
```


```{r}
label_SCP <- rep(NA, length(out$beta)-1)

sub1 <- tau[1:1]
label_SCP[sub1] <- 'Late'

sub2 <- tau[2:15]
label_SCP[sub2] <- 'Mid'

sub4 <- tau[16:length(mut_list)]
label_SCP[sub4] <- 'Early'
label_SCP <- c(NA, label_SCP)
```

```{r}
# clustering results from Pierre
dt_YTA <- read_xlsx('Table_Classes_TP53variants.xlsx')

label_YTA <- rep(NA, length(mut_list))
for (i in 1:length(mut_list)) {
  ind <- which(dt_YTA$ProtDescription==mut_list[i])
  if(length(ind)==1){
    label_YTA[i] <- dt_YTA$cluster_rec[ind]
  }
  
  if(grepl("*", mut_list[i], fixed = TRUE)){
    label_YTA[i] <- 0
  }
  if(grepl("fs", mut_list[i], fixed = TRUE)){
    label_YTA[i] <- 0
  }
  if(grepl("del", mut_list[i], fixed = TRUE)){
    label_YTA[i] <- 0
  }
  if(grepl("insIQ", mut_list[i], fixed = TRUE)){
    label_YTA[i] <- 0
  }
}
table(label_YTA)
```

```{r}
SCP_results <- data.frame(names = rownames(fit$beta), coef=Beta_matrix[, idx], label_SCP=label_SCP, label_YTA = c(NA, label_YTA))
df_ordered <- SCP_results[order(SCP_results$coef,decreasing=FALSE, na.last=FALSE), ]
print(df_ordered)
```
Now we construct the dataset for the KM plot. 


```{r}
dt_LFS_KM <- dt_LFS_use
dt_LFS_KM$Group <- NA

for (i in 1:nrow(dt_LFS_KM)) {
  mut <- dt_LFS_KM$ProtDescription[i]
  ind <- which(SCP_results$names %in% mut)
  dt_LFS_KM$Group[i] <- SCP_results$label_SCP[ind]
}

```


Now we are ready to show the KM plot
```{r}
# Load the libraries
library(survival)
library(survminer)
surv_object <- Surv(time = dt_LFS_KM$time, event = dt_LFS_KM$status)

# Fit a Kaplan-Meier survival curve
km_fit <- survfit(surv_object ~ Group, data = dt_LFS_KM)

# Plot the Kaplan-Meier curve
ggsurvplot(km_fit,
           data = dt_LFS_KM,
           pval = TRUE,             # Show p-value for the log-rank test
           conf.int = TRUE,         # Add confidence intervals
           risk.table = TRUE,       # Add risk table
           legend.labs = c("Early", "Mid", "Late"),  # Label the legend
           xlab = "Age at first cancer diagnosis",    # Label for x-axis
           ylab = "Survival Probability", # Label for y-axis
           ggtheme = theme_minimal() ,# Use a minimal theme
           font.main = c(16, "bold"),
           font.x = c(14), #, "italic"),
           font.y = c(14), #, "italic"),
           font.legend = c(14), #, "italic"),
           font.tickslab = c(12, "plain")
)

```

## Trans-SCP for LFS 

We now run Trans-SCP by using the additional survival information from the TP53 germline carriers in the TP53 database. 

### Run the Cox model to get the ranking from the source dataset.

```{r}
# sample size in the source dataset
n_S <- nrow(dt_Germ_use)

### Covariate matrix
X_S <- matrix(0, nrow = n_S, ncol = length(mut_list)+1)
colnames(X_S) <- c('gender', mut_list)
# covariates include gender + mutations
for (i in 1:n_S) {
  # put gender into the first column
  X_S[i, 1] <-  ifelse(dt_Germ_use$Sex[i]=='M', 1, 0) # gender. Male is 1, female is 0 
  # get the corresponding mutations 
  ind <- which(colnames(X_S)==dt_Germ_use$ProtDescription[i])
  X_S[i, ind] <- 1
}

### Survival information
Y_S <- cbind(time = dt_Germ_use$time, status = dt_Germ_use$status) 

## Preliminary estimation from the source
fit_S <- glmnet(X_S, Y_S, family = "cox", alpha = 0, lambda = 1e-4)
beta_S <- as.vector(fit_S$beta[-1]) 
```



```{r}
## Preliminary estimation from the target
fit_T <- glmnet(X_T, Y_T, family = "cox", alpha = 0, lambda = 1e-4)
beta_T <- as.vector(fit_T$beta[-1]) # we do not include the gender

# We use the number of carries to serve as weights
tb_S <- table(dt_Germ_use$ProtDescription)
tb_T <- table(dt_LFS_use$ProtDescription)

TB_S <- data.frame(name=rownames(tb_S), count=as.vector(table(dt_Germ_use$ProtDescription)))
TB_S <- TB_S[match(mut_list, rownames(tb_S)), ]
count_S <- TB_S$count


TB_T <- data.frame(name=rownames(tb_T), count=as.vector(table(dt_LFS_use$ProtDescription)))
TB_T <- TB_T[match(mut_list, rownames(tb_T)), ]
count_T <- TB_T$count


W_S <- count_S/(count_T+count_S)
W_T <- count_T/(count_T+count_S)

# calculate the weighted ranking 
tau_weighted <- wRankAveg(beta_S, beta_T, W_S, W_T)
```


Next, we are ready to run trans-SCP using the weighted ranking.

```{r}
# Construct the differencing matrix using the weighted ranking
X2 <- X_T[, -1]
X2_order <- X2[, tau_weighted]
X2_order <- cbind(X_T[, 1], X2_order)
D_weighted <- myGetDkn(k=1, n=ncol(X2_order)-1)
D_weighted <- cbind(rep(0, nrow(D_weighted)), D_weighted)

##########

ncSeq <- 1:5
nC <- length(ncSeq)
llkSeq2 <- rep(NA, nC)
bicSeq2 <- rep(NA, nC)
aicSeq2 <- rep(NA, nC)
nuSeq2 <- rep(NA, nC)
Beta_matrix2 <- matrix(NA, ncol(X_T), nC)

for (j in 1:nC) {
  cat(paste('Current stage:', j, 'case'))
  # fix the number of clusters
  nclusters <- ncSeq[j]
  

  t0 <- Sys.time()
  out_trSCP <- pcox(Y_T[, 1], Y_T[, 2], X2_order, D_weighted, penalty="distance",lambda=0, k=nclusters-1, 
                                    #beta0 = rnorm(ncol(X_order), 0,1),
                                     #beta0 = c(fit_T$beta[1], sort(beta0)),
                                    maxiters=1e3, tol=1e-5)
  t2 <- Sys.time()
  t2-t0

  beta_hat_trSCP <- beta_order_trSCP <- out_trSCP$beta[-1]
  for(i in 1:(length(beta_order_trSCP))){
      beta_hat_trSCP[tau_weighted[i]] <- beta_order_trSCP[i]
  }
  beta_hat_trSCP <- c(out_trSCP$beta[1], beta_hat_trSCP)
  Beta_matrix2[, j] <- beta_hat_trSCP 

  
  llkSeq2[j] <- pll_cox(Y_T[, 1], Y_T[, 2], X_T, beta_hat_trSCP)

  nu <- nuSeq2[j] <-  length(which(abs(D_weighted%*%out_trSCP$beta)>1e-2))
  #tau <- choose(nrow(D_weighted), nu)
  bicSeq2[j] <- -2*llkSeq2[j] + log(nrow(X2_order))*nu
  aicSeq2[j] <- -2*llkSeq2[j] + 2*nu
  
}

save.image("TP53_test_final.RData")
```

```{r}
idx <- which.min(bicSeq2)

st_trSCP <- sort(Beta_matrix2[-1, idx], index.return = TRUE)
plot(st_trSCP$x)

which(diff(st_trSCP$x)>0.05)
```


```{r}
label_trSCP <- rep(NA, length(out_trSCP$beta)-1)

Loc <- st_trSCP$ix
sub1 <- tau_weighted[1:4]
label_trSCP[sub1] <- 'Late'

sub2 <- tau_weighted[5:16]
label_trSCP[sub2] <- 'Mid'

sub4 <- tau_weighted[17:length(mut_list)]
label_trSCP[sub4] <- 'Early'
label_trSCP <- c(NA, label_trSCP)
```

```{r}
trSCP_results <- data.frame(names = rownames(fit_T$beta), coef=Beta_matrix2[, idx], label_trSCP=label_trSCP, label_YTA = c(NA, label_YTA))
trSCP_ordered <- trSCP_results[order(trSCP_results$coef, decreasing=FALSE, na.last=FALSE), ]
print(trSCP_ordered)
```

Now Let us construct the dataset for the KM plot. 

```{r}
dt_LFS_KM2 <- dt_LFS_use
dt_LFS_KM2$Group <- NA

for (i in 1:nrow(dt_LFS_KM2)) {
  mut <- dt_LFS_KM2$ProtDescription[i]
  ind <- which(trSCP_results$names %in% mut)
  dt_LFS_KM2$Group[i] <- trSCP_results$label_trSCP[ind]
}

```


Now we are ready to show the new KM plot
```{r}
# Load the libraries
library(survival)
library(survminer)
surv_object <- Surv(time = dt_LFS_KM2$time, event = dt_LFS_KM2$status)

# Fit a Kaplan-Meier survival curve
km_fit <- survfit(surv_object ~ Group, data = dt_LFS_KM2)

# Plot the Kaplan-Meier curve
 ggsurvplot(km_fit,
           data = dt_LFS_KM2,
           pval = TRUE,             # Show p-value for the log-rank test
           #conf.int = TRUE,         # Add confidence intervals
           #risk.table = TRUE,       # Add risk table
           legend.labs = c("early-onset", "late-onset", "mid-onset"),  # Label the legend
           xlab = "Age at first cancer diagnosis",    # Label for x-axis
           ylab = "Survival Probability", # Label for y-axis
           ggtheme = theme_minimal() ,# Use a minimal theme
           font.main = c(16, "bold"),
           font.x = c(14), #, "italic"),
           font.y = c(14), #, "italic"),
           font.legend = c(14), #, "italic"),
           font.tickslab = c(12, "plain")
)

```


```{r}
Combined_results <- data.frame(names = rownames(fit_T$beta), coef_SCP=Beta_matrix[, idx], coef_trSCP=Beta_matrix2[, idx],
                               label_SCP=label_SCP,label_trSCP=label_trSCP, label_YTA = c(NA, label_YTA), 
                               count = c(NA, TB_T$count))
print(Combined_results)

# save the results 
library(writexl)
write_xlsx(Combined_results, 'TrSCP_combined_final2.xlsx')
```

The KM plot using YTA analysis.

```{r}
dt_LFS_KM3 <- dt_LFS_use
dt_LFS_KM3$Group <- NA

for (i in 1:nrow(dt_LFS_KM3)) {
  mut <- dt_LFS_KM3$ProtDescription[i]
  ind <- which(dt_YTA$ProtDescription %in% mut)
  if(length(ind)==1){
      if(dt_YTA$cluster_rec[ind] %in% c("0", "A")){
         dt_LFS_KM3$Group[i] <- "High"
      }
    
      if(dt_YTA$cluster_rec[ind] %in% c("B")){
         dt_LFS_KM3$Group[i] <- "Medium"
      }
    
      if(dt_YTA$cluster_rec[ind] %in% c("C", "D")){
         dt_LFS_KM3$Group[i] <- "Low"
      }
  }
  if(grepl("*", mut, fixed = TRUE)){
    dt_LFS_KM3$Group[i] <- "High"
  }
  if(grepl("fs", mut, fixed = TRUE)){
    dt_LFS_KM3$Group[i] <- "High"
  }
  if(grepl("del", mut, fixed = TRUE)){
    dt_LFS_KM3$Group[i] <- "High"
  }
  if(grepl("insIQ", mut, fixed = TRUE)){
    dt_LFS_KM3$Group[i] <- "High"
  }
  

}

surv_object <- Surv(time = dt_LFS_KM3$time, event = dt_LFS_KM3$status)

# Fit a Kaplan-Meier survival curve
km_fit <- survfit(surv_object ~ Group, data = dt_LFS_KM3)

# Plot the Kaplan-Meier curve
ggsurvplot(km_fit,
           data = dt_LFS_KM3,
           pval = TRUE,             # Show p-value for the log-rank test
           #conf.int = TRUE,         # Add confidence intervals
           #risk.table = TRUE,       # Add risk table
           legend.labs = c("High-risk (0+A)", "Low-risk (c+D)", "Medium-risk (B)"),  # Label the legend
           xlab = "Age at diagnosis",    # Label for x-axis
           ylab = "Survival Probability", # Label for y-axis
           ggtheme = theme_minimal(), # Use a minimal theme
           font.main = c(16, "bold"),
           font.x = c(14), #, "italic"),
           font.y = c(14), #, "italic"),
           font.legend = c(14), #, "italic"),
           font.tickslab = c(12, "plain")
)
```

